Efficient, Robust and Exact Boolean Operations on N Triangular Mesh Primitives

### 2016-07-14 summary.tex
tess-graph

\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
%\usepackage{ulem}
%\usepackage[numbers,sort&compress]{natbib}
\renewcommand{\arraystretch}{1.3}

@@

The concept of plane-based representation of polygonal meshes was first described by Sugihara and Iri \cite{sugihara1990solid}. Plane-based representation provides one important advantage
when it comes to tasks that involve changing the topology of solids represented by meshes like the evaluation of Boolean expressions: no new primary geometry information has to be constructed to obtain the resulting polyhedron – it is composed of a subset of the planes of the input polyhedra. Hence, opposed to the case of using vertex coordinates, that inevitably necessitates the construction of new geometry information, only geometric decision predicates are required to compose the output polyhedron from the face planes of the input. Since the input is usually given in a numerical representation with finite precision, we can determine an upper bound on the precision that is needed to make correct decisions. The a priori knowledge of this upper bound allows us to use fixed precision predicates that are specifically tailored to the precision required in the worst case, resulting in a vastly better performance compared to techniques for arbitrary numerical precision, that are usually required when
constructions are part of the processing. Our processing is rigorously based on this paradigm of plane-based geometry representation that allows us to perform fully robust, exact computations using only fast fixed precision predicates. These predicates take planes as arguments and check coincidence and co-orientation of planes, orientation of a plane with respect to a point defined by the intersection of three planes, and whether three planes intersect in a unique point. The latter one can also be used in negated form to check if three planes, that are known to intersect in a common point, even intersect in a common line. We implement them using filtering techniques proposed by Shewchuk.

@@

Once all intersections are detected, we need to tessellate the input primitives and construct the Linked Halfedges.. In many methods like \cite{ogayar2015deferred}, constraint Delaunay triangulation (CDT) is applied to perform tessellation, since the intersections are naturally the constraints of CDT. However, for a CSG with more than two primitives, the intersection vertices may be generated by intersections of three faces from different primitives, which cannot be computed during the previous step. Also, intersection line segments may overlap or intersect with each other, and cannot be used for constraints of CDT directly. Finally, as our intersection vertices are represented exactly by planes, implementation of CDT could be complex and inefficient, as most CDT methods were not designed to handle planes. Therefore, we design a simple structure, called tessellation graph, to guide the tessellation of each single face, and develop a very fast but exact method to construct Linked Halfedges from tessellation graphs, which is suitable for dealing with plane-based representation of intersections. Details are shown in section \ref{sec:tessellation}.

@@

\section{Plane-based geometry}

@@

\subsection{Vertex-based method review}

@@

\subsection{Plane-based technique embedding}

@@

\subsection{Intersection record}
\label{sec:ir}

@@

We recall our definition of Linked Halfedge in section \ref{sec:meshes}, we need to transfer necessary information by IR to tessellation stage for Linked Halfedge construction. For each intersection line segment, we use a five-component vector  $\gamma \colon(v_b, v_e, f, p, C)$ to describe it. Here, $v_b$ and $v_e$ are the two end points of $\gamma$. Component $f$ is the face where the $\gamma$ lies. Component $C$ is the intersection context, and will be further explained later. Component $p$ is the plane representation of the line where $\gamma$ lies (denoted as $\vec{n}_\gamma$). Normally, line is represented by two planes. However, one of the planes for $\vec{n}_\gamma$ must be $f$'s supporting plane $p^s(f)$, so we just omit it, and use the other plane as the representation for $\vec{n}_\gamma$ .

@@

\subsection{Degenerate cases}
\label{sec:degenerate}

@@

\subsubsection{Intersects on face edges}
When intersection line segment lies on an edge (Fig. \ref{fig:twin}). we call it as \emph{edge intersections}. There are generally two kinds of edge intersections. If $\gamma$ lies on the edge of $f_1$ or $f_2$, but inside of the face of the other one, then we called it as edge-face intersections. If $\gamma$ lies on the edges of both $f_1$ and $f_2$, we call it as edge-edge intersections. When there is edge-face intersection, we treat it as the intersection of three faces $f_1$, $f_2$ and $f_1^\star$ (or $f_2^\star$). When there is edge-edge intersection, we treat it as the intersection of four faces $f_1$, $f_2$, $f_1^\star$ and $f_2^\star$. The $f_1^\star$ (or $f_2^\star$), known as \emph{companion face},  is the neighboring face which shares the intersection edge with $f_1$ (or $f_2$). In this way, we avoid repetitive insertion of the same edge intersections, since the companion faces share the same intersection with its neighborhood $f_1$ or $f_2$.

@@

When two triangle faces are coplanar, they may intersect with each other depending on whether they overlap in some area. Many methods \cite{feito2013fast} claim that it is important to compute the overlap in a 2D intersection way and to tessellate the face according to the intersection result afterwards. We will show that this is not necessary in our method. We simply omit coplanar cases in our intersection test and it will not affect the correctness of our tessellation stage.

@@

Consider two triangle faces $f_1$ and $f_2$ intersect within the supporting plane (Fig. \ref{fig:coplanar}). Both $f_1$ and $f_2$ will be divided into two areas---the overlap area and the exclusive part. The boundary of these two parts are the boundary from the other face. Following the concision principle, whether we have to split face according to these boundaries depends on the necessity. That is, we have to

@@

\label{sec:tessellation}
This is the step to process TTIR and output LHIC, where the tessellation happens. We called this process as deferred tessellation because we do not tessellate until we collect all intersection line segments. For this problem, Ogayar-Anguita et. al. \cite{ogayar2015deferred} try to use Constrained Delaunay Triangulation (CDT) to solve, as the recorded intersection is naturally the constraints. It works fine for Boolean operation of two primitives. However, when there are more than two primitives, simply using CDT covers the complexity of tessellation. In such case, the intersection line segments may cross, overlap and even coincident with each other, introducing new vertex, splitting original intersection line segments. Naive implementation of CDT may lead to holes, repetitive edges, and other topology errors.

To solve these problems, we first refine the TTIR into an intermediate data structure---\emph{tessellation graph}. In this process, we resolve any intersection between intersection. After this, we tessellation each face according its corresponding tessellation graph. To make this process unconditional robust and exact, we utilize plane-based techniques and do not compute explicitly the coordinates of intersection points throughout this step.

\subsection{Tessellation graph}

TTIRs are only loose data that store the coordinates of intersection line segments. To perform tessellation and extract sub-faces, we need to organize TTIRs in a face scope. We use a graph-based data structure---tessellation graph---for this purpose. For each intersection face which needs tessellation, we construct tessellation graph. A tessellation graph is the graph-like description of the tessellated face topology. Nodes of tessellation graph represent vertices, and edges represent edges in LHIC. To avoid ambiguity, we call original three edges of the face as \emph{face boundaries} in the following. An important feature of tessellation graph is that it uses uniform representation for intersection vertices and line segments, and face boundaries and the original three corners. This dramatically reduces complexity of extracting sub-faces and allowing the following process simple and fast.

\subsubsection{Definition and construction}

Nodes of tessellation graph represent vertices and store the coordinates of them. Edges are nondirectional ones, and store the intersection context if exist. We classify graph edges into two types: we call edges on face boundaries as \emph{boundary edges} and edges inside face as \emph{inner edges}. They are different in at least two aspects: first, the boundary edges may not have a intersection context since it is not coincident with any TTIR; second, in the view of halfedge, inner edges represent two halfedges but boundary edges represent only one halfedge. The same classification is for intersection line segments: if the intersection are inside the face, it is an inner intersection, otherwise a boundary intersection. //可以提前到degenreate case

Recall our definition of TTIR in section \ref{sec:ir}, we find most of the work has already been done. To construct inner edges, we only need to first collect all vertices on the face (including both inside vertices and boundary vertices) to be the nodes, and then connect nodes by inner intersections. However, boundary edge construction is not that easy. We cannot simply connect the three face corners to be the boundary edges, since there may be boundary intersections, and face edges have to be splitted.

@@

\subsubsection{Intersection refinement}

Since tessellation graph is a graph-like description of halfedge, it has extra constraints on its form. The most important one is that edges should not intersect each other. This can be guaranteed when there are only two primitives,  because the surface of primitives is all watertight, non-manifold, and without self-intersection. However, when there are more than two primitives, this constraint can be violated. In the view of intersection detection, it is not enough to find all intersection vertices by only triangle-triangle intersection test. This is because the new vertex can be formed not only by edge intersects plane, but also by three planes, which involves more than two triangles. Therefore, an extra processing--intersection refinement---is necessary to resolve these intersections before constructing tessellation graph.

Intersection refinement is performed between every two intersection line segments from different primitives. Using plane-based representation, whether two intersection line segments intersect is determined by predicate of orientation of plane against a point. For example, we want determine whether $I_v$: $(v_b, v_e, P_v, f_0, C_v)$ and $I_u$: $(u_b, u_e, P_u, f_0, C_u)$ intersects.
% 需要在IR那一章补充plane rep of intersection这一部分。intersection is describe with 四元组。
It is done by checking whether $u_b, u_e$ is on different sides of $P_v$ and whether $v_b, v_e$ is on different sides of $P_u$. If yes, $I_v$ and $I_u$ intersects and introduced new intersection vertex, which is the intersection of $P_u, P_v$ and the supporting plane $f_0$.

Degenerate cases will occur when $I_u$ and $I_v$ intersect on end point(s) or $I_v$ and $I_v$ are collinear.

\subsection{Extracting sub-faces}
The last step of tessellation is to really change the topology of input primitive to construct LHICs. We could apply CDT here since the intersection is refined and there is no intersection between the constraints. However, we choose to design a new method to extract sub-faces from intersection face for the efficiency:
\begin{itemize}
  \item lose topology information
  \item may not need to triangulate, classify in a whole
  \item hard to implement with planes
\end{itemize}

There is yet another problem on this method. Our method of extracting sub-faces assume that the tessellation graph is connected. However, this assumption can be false (Fig.).

@@

\subsection{Basic idea}

@@

\subsection{Fast and exact face classification}

@@

\subsection{Caching evaluation results}

>>>>

\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
\usepackage{color}
%\usepackage{ulem}
%\usepackage[numbers,sort&compress]{natbib}
\renewcommand{\arraystretch}{1.3}

@@

The concept of plane-based representation of polygonal meshes was first described by Sugihara and Iri \cite{sugihara1990solid}. Plane-based representation provides one important advantage
when it comes to tasks that involve changing the topology of solids represented by meshes like the evaluation of Boolean expressions: no new primary geometry information has to be constructed to obtain the resulting polyhedron ¨C it is composed of a subset of the planes of the input polyhedra. Hence, opposed to the case of using vertex coordinates, that inevitably necessitates the construction of new geometry information, only geometric decision predicates are required to compose the output polyhedron from the face planes of the input. Since the input is usually given in a numerical representation with finite precision, we can determine an upper bound on the precision that is needed to make correct decisions. The a priori knowledge of this upper bound allows us to use fixed precision predicates that are specifically tailored to the precision required in the worst case, resulting in a vastly better performance compared to techniques for arbitrary numerical precision, that are usually required when
constructions are part of the processing. Our processing is rigorously based on this paradigm of plane-based geometry representation that allows us to perform fully robust, exact computations using only fast fixed precision predicates. These predicates take planes as arguments and check coincidence and co-orientation of planes, orientation of a plane with respect to a point defined by the intersection of three planes, and whether three planes intersect in a unique point. The latter one can also be used in negated form to check if three planes, that are known to intersect in a common point, even intersect in a common line. We implement them using filtering techniques proposed by Shewchuk.

@@

Once all intersections are detected, we need to tessellate the input primitives and construct the Linked Halfedges.. In many methods like \cite{ogayar2015deferred}, constraint Delaunay triangulation (CDT) is applied to perform tessellation, since the intersections are naturally the constraints of CDT. However, for a CSG with more than two primitives, the intersection vertices may be generated by intersections of three faces from different primitives, which cannot be computed during the previous step. Also, intersection line segments may overlap or intersect with each other, and cannot be used for constraints of CDT directly. Finally, as our intersection vertices are represented exactly by planes, implementation of CDT could be complex and inefficient, as most CDT methods were not designed to handle planes. Therefore, we design a simple structure, called Tess-graph, to guide the tessellation of each single face, and develop a very fast but exact method to construct Linked Halfedges from Tess-graphs, which is suitable for dealing with plane-based representation of intersections. Details are shown in section \ref{sec:tessellation}.

@@

\section{Plane-based Geometry}

@@

\subsection{Vertex-Based Method Review}

@@

\subsection{Plane-based Technique Embedding}

@@

\subsection{Intersection Representation}
\label{sec:ir}

// change the title to Indentifying intersections

@@

We recall our definition of Linked Halfedge in section \ref{sec:meshes}, we need to transfer necessary information by IR to tessellation stage for Linked Halfedge construction. For each intersection line segment, we use a five-component vector {\color{red}{$\gamma \colon(T, v_0, v_1, p, C)$}} to describe it. Here, $v_b$ and $v_e$ are the two end points of $\gamma$. Component $f$ is the face where the $\gamma$ lies. Component $C$ is the intersection context, and will be further explained later. Component $p$ is the plane representation of the line where $\gamma$ lies (denoted as $\vec{n}_\gamma$). Normally, line is represented by two planes. However, one of the planes for $\vec{n}_\gamma$ must be $f$'s supporting plane $p^s(f)$, so we just omit it, and use the other plane as the representation for $\vec{n}_\gamma$ .

@@

\subsection{Degenerate Cases}
\label{sec:degenerate}

@@

\subsubsection{Intersects on face edges}

// intersect on face edges does not create new line segments, it is part of the edges from other meshes, and will to be duplicated.

// whose edge? need to be further explained and better defined. an edge intersection divide edge in a triangle and will be duplicated in the other

When intersection line segment lies on an edge (Fig. \ref{fig:twin}). we call it as \emph{edge intersections}. There are generally two kinds of edge intersections. If $\gamma$ lies on the edge of $f_1$ or $f_2$, but inside of the face of the other one, then we called it as edge-face intersections. If $\gamma$ lies on the edges of both $f_1$ and $f_2$, we call it as edge-edge intersections. When there is edge-face intersection, we treat it as the intersection of three faces $f_1$, $f_2$ and $f_1^\star$ (or $f_2^\star$). When there is edge-edge intersection, we treat it as the intersection of four faces $f_1$, $f_2$, $f_1^\star$ and $f_2^\star$. The $f_1^\star$ (or $f_2^\star$), known as \emph{companion face},  is the neighboring face which shares the intersection edge with $f_1$ (or $f_2$). In this way, we avoid repetitive insertion of the same edge intersections, since the companion faces share the same intersection with its neighborhood $f_1$ or $f_2$.

@@

@@

Consider two triangle faces $T_1$ and $T_2$ intersect within a common plane. Both $T_1$ and $T_2$ will divide each other into two areas---the overlapped area and the exclusive part (Fig. \ref{fig:coplanar}). Obviously, if we tessellate both $T_1$ and $T_2$ according to the boundary of overlapped area, we can guarantee that the tessellated meshes will not cross the boundary of any other meshes. In fact, many methods \cite{feito2013fast} claim that it is important to tessellate faces according to the 2D-intersection result. We will show that this is not necessary in our method. We do not care whether $T_1$ and $T_2$ really intersect in 2D. We simply omit all coplanar cases, treating them as they do not intersect. This simplify our intersection test, making it more robust and fast (coplanar cases can be usual in a CSG of artificial object), while doing no harm to the correctness of our method.

The key observation is that coplanar cases do not create new edges, just like edge intersection in the last section. The difference is that in a non-coplanar case, there is only one edge intersection. In a coplanar, there are at most three intersections. Therefore, we can decompose it into three edge intersections and record each edge intersection independently. However, there is a better way that we do not process them at all. This is because there is companion face, like $T_1^*$, in edge intersection. If $T_1^*$ is not coplanar with $T_2$, then we can solve the intersection between $T_1^*$ and $T_2$ in a 3D way, which is the case in Fig \ref{fig:twin}(c). On the other hand, if $T_1^*$ is coplanar with $T_2$, we instantly know that both sides of shared edge $e(T_1, T_1^*)$ have the same indicator $\lambda_1$, which imply that this intersection is not necessary for tessellation and we can omit it by the principle of concision. In sum-up, we do not need to handle with coplanar intersection cases.

@@

{\color{red}{red alert start}}

\label{sec:tessellation}
This is the step to tessellate input meshes with intersections, outputting linked halfedge. We called this stage as deferred tessellation because it happens until we collect all necessary intersections. For this problem, Ogayar-Anguita et. al. \cite{ogayar2015deferred} try to use Constrained Delaunay Triangulation (CDT) to solve, as the intersections is naturally the constraints and triangle faces is naturally a convex zone. It works fine for Boolean operation of two primitives. However, when there are more than two primitives, simply applying CDT covers the complexity of tessellation. The major problem here is that the intersections may be \emph{irregular}. The definition of regular intersection is that it does not cross or overlap with other intersections. In CSG cases with more than two meshes, irregular intersections can introduce new vertex, split original intersection line segments. Naive algorithm of CDT [there should be algorithm for instance] may lead to holes, repetitive edges, and other topology errors when there are irregular intersections {\color{red}{(Fig. ?, use CGAL or fade2d for testing)}}.

To solve these problems, we first refine the intersections and eliminate any irregular intersections. When all intersections are regular, we can directly use CDT to tessellate. {\color{red}{However?xxxx(after I research CDT)}}. Therefore, we develop a new tessellation algorithm that guarantee topology correctness and unconditional robust. We organize intersections into single face scope, generating an intermediate data structure---\emph{Tess-graph}, After that, we tessellate each face by subtract sub-faces from its corresponding Tess-graph. Throughout this stage, we utilize plane-based techniques and do not compute explicitly the coordinates of intersection points.

\subsection{Eliminating Irregular Intersection}

Regular intersections are guaranteed when there are only two meshes, because the surfaces of primitives are all without self-intersection. However, when there are more than two primitives, this constraint can be violated. In the view of intersection detection, it is not enough to find all intersection vertices by only triangle-triangle intersection test. This is because in such case, the new vertex can be introduced not only by intersection between edge and plane, but also by intersections of three planes, which involves more than two faces.

Intersection refinement is performed in a local scope. For each face, we detect irregular intersections by intersection test between pairs of intersections. Because primitive meshes are not self-intersected, we only check intersections between intersections made by different meshes. Face intersecting only one other mesh do not need refinement, since they do not have such intersection pairs.

Using plane-based representation, whether two intersections intersects each other is determined by predicate of orientation of plane against a point. For example, we want determine whether $\lambda_1$ and $\lambda_2$ intersects. Using the notation of section \ref{sec:ir}, we perform the test by checking whether $v_0(\lambda_1), v_1(\lambda_1)$ are on two sides of $p(\lambda_2)$ and whether $v_0(\lambda_2), v_1(\lambda_2)$ are on two sides of $p(\lambda_1)$. If yes, $\lambda_1$ and $\lambda_2$ intersects and introduces new intersection vertex $v_{new}$, which is the intersection of $p(\lambda_1), p(\lambda_2)$ and the supporting plane of $T(\lambda_1)$. The new vertex $v_{new}$ should be added into both $\lambda_1$ and $\lambda_2$ as a segment point. Since there may be more than one segment point added into a certain intersection during the whole test, we do not divide the intersection until all intersection pairs are tested.

Degenerate cases are when the two intersections intersect on one of the end points or are collinear. In the first condition {\color{red}{Fig. ??(a)}}, there is no new vertex generated and only one of the intersection needs splitting. The collinear case {\color{red}{Fig. ??(b)}} is intersection in 1D and are tested exactly by two-point-on-line comparison between the end points of intersections. If there is intersection between them, it should be straightforward to figure out the segment points for each intersections.

After all segment points are found, the related intersections have to be split. For a certain intersection, we sort the segment points by two-point-on-line comparisons and subtract child intersections along the direction of their father one by one. The representation of the child intersections should be the same with their father except the coordinate components $v_0$ and $v_1$. So far, we guarantee that all refined intersections need not be further divided as the potential edges of the Linked Halfedge. However, if there are collinear cases, there will be coincident child intersections from different fathers. We choose to resolve them in Tess-graph. If CDT is chosen instead of Tess-graph, coincidence can be handled like-wise.

\subsection{Tess-Graph}

So far, intersections are only loose data that store the coordinates and neighbouring information. To perform tessellation and extract sub-faces, we need to organize them to reveal the topology. We use Tess-graph for this purpose. A Tess-graph is the graph description of the tessellated face topology. For each face to be tessellated, we construct a Tess-graph according to the refined intersections. The final Linked Halfedge merges both the original topology and the intersections in the same representation. And the construction of Tess-graph is the first step towards such uniformity. By only nodes and connections, it represents both the intersections and the original boundary of triangle face. This reduces complexity of the following processes and allowing them to be simple and fast.

Nodes of Tess-graph represent vertices and store the coordinates of them. Connections are nondirectional, represent edges in Linked Halfedge, storing the I-rep if it is coincident with an intersection. Sometimes, there will be more than one coincident intersections and all I-reps have to be stored.

We classify connections into two types: we call connections on face boundaries as \emph{boundary connections} and connections inside face as \emph{inner connections}. They are different in two aspects: first, the boundary connections may not have an intersection representation since it is not coincident with any intersection; second, in the view of halfedge, inner connections represent two halfedges but boundary connections represent only one halfedge. {\color{red}{The same classification is for intersection line segments: if the intersection are inside the face, it is an inner intersection, otherwise a boundary intersection. // can be forward to the previous place}}

Construction of Tess-graph from I-rep is straightforward except one thing: we have to merge face boundary with the intersections lies on the boundary.

Using the notation of I-rep, we find most of the work has already been done. To construct inner edges, we only need to first collect all vertices on the face (including both inside vertices and boundary vertices) to be the nodes, and then connect nodes by inner intersections. However, boundary edge construction is not that easy. We cannot simply connect the three face corners to be the boundary edges, since there may be boundary intersections, and face edges have to be splitted.

@@

\subsection{Extracting Sub-faces}
The last step of tessellation is to really change the topology of input primitive to construct LHICs. We could apply CDT here since the intersection is refined and there is no intersection between the constraints. However, we choose to design a new method to extract sub-faces from intersection face for the efficiency:
\begin{itemize}
  \item lose topology information
  \item may not need to triangulate, classify in a whole
  \item hard to implement with planes
\end{itemize}

There is yet another problem on this method. Our method of extracting sub-faces assume that the Tess-graph is connected. However, this assumption can be false (Fig.).

@@

\subsection{Basic Idea}

@@

\subsection{Fast and Exact Face Classification}

@@

\subsection{Caching Evaluation Results}
